Exploring the Impact of Encouragement and Student Behaviour on Academic Effort in Second-year Statistics Courses

### 2024-12-04 paper/paper.qmd
Update "paper" with modifying section title
update both "paper.qmd" and "paper.pdf"
Update the "paper.qmd" and "paper.pdf"
fixed some format problem, grammar problem, typos, and changed the subtitle

---
title: "Exploring the Impact of Encouragement and Student Behaviour on Academic Effort in Second-year Statistics Courses"
subtitle: "Motivational Emails Boost Effort, Perception Drive Effort, Difficulty Reduces Effort"
author: 
  - Shuangyuan Yang
thanks: "Code and data are available at: [https://github.com/Sophiaaa-Y/student-effort-analysis](https://github.com/Sophiaaa-Y/student-effort-analysis.git)."
date: today
date-format: long
abstract: "Student participation is important for academic success. This research employs Bayesian multiple linear regression to examine the effects of task complexity, intrinsic motivation, emotional states, cognitive competence, and types of email interventions on student effort in email-based programs. The findings emphasize the necessity of balancing task difficulty to prevent demotivation, while emotional and motivational factors, especially changes in cognitive competence and interest, substantially increase effort. The findings provide practical advice for developing personalized, economical strategies that combine cognitive and emotional support to enhance academic motivation."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

This study examines the variables affecting how students' study efforts changed in response to an academic intervention sent by email. The intervention's goal was to motivate students to become more interested in statistics and to actively learn about it. The study examines factors like emotional feelings, cognitive competence, perceived difficulty, intrinsic motivation, and email type in order to determine how these determinants all work together to influence differences in effort. The analysis uses more than 600 student records and a Bayesian multiple linear regression model generated by the `rstanarm` package in R [@rstanarm] to determine how these factors together affect academic engagement. In contrast to other research that looked at these factors separately [@akbari2020students] [@goes2016incentive], this study provides a thorough framework for understanding how different facets of student engagement and strategies for communication interact to influence behavioral changes under academic background.

@@

## Data Overview

@@

The dataset used in this study was sourced from the Nudge Data Repository [@Taback_nudgedata_2022]. It reflects real-world educational phenomena, examining how email interventions influence student attitudes toward statistic. Each observation corresponds to a unique student who received an intervention, with data collected on their behaviors and performance metrics before and after the emails. According to @taback2023randomized, the measurements in this study were carefully designed to align with established behavioral and educational theories, ensuring reliability and validity. Self-reported surveys were used to capture subjective variables, such as changes in emotional states, cognitive competence, and intrinsic motivation. These surveys employed structured Likert-scale items, ranging from "strongly disagree" to "strongly agree," allowing participants to express complex degrees of agreement or disagreement. Emotional states, such as interest and frustration, were assessed before and after the intervention to evaluate the effectiveness of motivational strategies, while cognitive competence was measured by asking participants to rate their confidence in handling statistical tasks. Objective metrics complemented these surveys by providing unbiased data on task complexity and performance outcomes. Task complexity was quantified based on the number of steps required to complete problems and the time spent on each task, while performance outcomes were measured against standardized benchmarks, such as quiz scores and task completion rates. This dual-method approach ensured a thorough understanding of how interventions influenced behavior, capturing both psychological engagement and actual performance. Grounded in theoretical models like self-determination theory and cognitive load theory, the study integrated subjective and objective measures to provide a robust framework for analyzing the predictors of academic effort and engagement.

@@

- **change in affect:** Change in students’ feelings concerning to the course material before and after the intervention. Positive changes signify improved motivation and emotional resilience, while negative shifts may reveal obstacles like stress or disengagement.
- **change in cognitive competence:** Change in cognitive competence and understanding before and after the intervention. Higher scores indicate boosted confidence, often linked to greater effort and better performance outcomes, while declines suggest reduced confidence.
- **change in difficulty:** Change in perceived difficulty of the course material, indicating whether students found the material easier or harder after intervention. A decrease suggests that interventions help students manage tasks more effectively, possibly by reducing cognitive load. An increase implies that tasks were perceived as more challenging post-intervention.
- **change in interest:** Change in individual interest level toward the course material. Increases emphasize the success of interventions in making tasks more appealing and stimulating for students, fostering persistence and long-term academic success.
- **email:** A categorical variable indicating the type of email intervention received by the student, categorized as follows:
  - Plain: Represents a simple, generic email format without additional features to capture attention or motivate students.
  - Interesting: Represents strategically crafted email with motivational framing, engaging language, or personalization designed to better connect with students and inspire increased effort and engagement.

@@

- **change in affect:** Tracks shifts in students' emotional states pre-intervention and post-intervention, showing how interventions influence feelings about academic tasks. Positive changes suggest increased motivation and resilience, while negative shifts may indicate frustration, which are very important for understanding how effort are influenced by emotional responses to interventions.
- **change in cognitive competence:** This means change in cognitive competence, which measures changes in students’ perceived learning ability, providing useful information about self-efficacy and confidence. Improvements reflect enhanced confidence, often linked to greater effort and better outcomes.
- **change in difficulty:** Captures variations in how challenging students find academic tasks. Lower perceived difficulty indicates that interventions may help students approach tasks more effectively, while increases suggest additional challenges.
- **change in interest:** Evaluates changes in intrinsic motivation and engagement, which are important for encouraging long-term academic persistence. Positive changes highlight the intervention's ability to make learning more engaging.
- **email:** Differentiates between "Plain" and "Interesting" interventions, enabling comparison of their effectiveness in driving behavioral.

@@

The dataset's structure and the goal of the study, which is to identify the variables impacting shifts in student effort, serve as the foundation for the modeling choices. As they indicate particular behavioral or attitudinal changes after the intervention and represent important characteristics determining effort, "Change in Affect", "Change in Cognitive Competence", "Change in Difficulty", and "Change in Interest" are included as predictor variables. To maintain precision and prevent information loss, these variables are considered as continuous. besides, to explore the different impacts of the intervention method on effort, another predictor variable, "Email type", is included as a binary independent variable, with values of "Plain" and "Interesting". Pre- and post-intervention baseline variables are not included since the predictor variables have already reflected changes, which prevents redundancy. This approach aligns with the dataset's structure and the hypothesis that shifts in emotional states, perceived competence, task difficulty, intrinsic interest, and email type, collectively influence effort.

@@

kable(model_summary_df, format = "markdown", align = "c", col.names = c("Parameter", "Mean", "SD", "10%", "50%", "90%"))
```
@tbl-model-summary provides the model's coefficient summary, which shows the estimated impacts of predictors on the outcome variable "Change in Effort." With its posterior mean, standard deviation (SD), and credible intervals at 10%, 50%, and 90%, each measure provides a clear picture of the variables affecting student effort after the intervention.

- Change in Affect, which has mean equals to 0.25 and standard deviation equals to 0.10. This parameter reflects the power of emotions in shaping behaviors. When students experience shifts in their emotional state, perhaps feeling more positive or encouraged, their effort increases. This effect is not only positive but also reliably significant, as shown by the credible interval from 0.10 to 0.40. Such results suggest that fostering a supportive emotional environment can play an important role in encouraging students to participate more actively in their tasks.

- Change in Cognitive Competence, which has mean equals to 0.12 and standard deviation equals to 0.08, emphasizes the impact of students' self-perception of learning ability. When students feel more confident in their cognitive skills, their effort also rises, even though it's not as significant as emotional changes. The credible interval from 0.01 to 0.23 illustrates the more complex aspect of this impact, demonstrating that although increases in self-perceived competence are important, they often take second place to other factors like interest changes.

- Change in Difficulty Perception, which has mean equals to -0.05 and standard deviation equals to 0.07. There is a minor tendency for effort to decrease as tasks are considered as becoming more challenging. When presented with more difficult tasks, some students may feel overwhelmed or frustrated, according to this effect, which is represented by a little negative mean value. There is diversity, too, as the credible interval from -0.20 to 0.10 shows that some kids may respond to challenges with less effort, while others may rise to the challenge and show resilience. This variation emphasizes the various ways people react to perceived difficulties, demonstrating the necessity of individualized intervention strategies.

- Change in Interest, which has mean equals to 0.30 and standard deviation equals to 0.11. This predictor emphasizes the significance of intrinsic drive by showing the strongest positive effect. Students' work significantly increases when they find a task interesting or engaging. Raising interest can be one of the best strategies to improve student engagement, as the credible interval that range from 0.10 to 0.50 illustrates to the effect's stability. These findings demonstrate that student-centered interventions have the potential to be transformative in addition to having an impact.

- Email Type: Interesting, which has mean equals to 0.15 and standard deviation equals to 0.10. Although it has a less consistent effect, email type also shows a favorable impact. The reasonable interval from 0.00 to 0.30 indicates that creating captivating, attention-grabbing emails seems to need more work. Though some students may benefit significantly from this customized communication, others might need more or other tactics to be completely motivated, as indicated by the wide range of responses.

@@

Additionally, these selected columns were renamed for clarity and interpretability. For example, "change_affect" was renamed to "change in affect", "change_cog" was renamed to "change in cognitive competence", "change_difficulty" was renamed to "change in difficulty", "change_interest" was renamed to "change in interest", and "change_effort" was renamed to "change in effort". These renaming steps improved the dataset's alignment with the study’s focus on behavioral outcomes.

@@

@tbl-model-results presents the regression results for the "Change Effort Model," outlining the estimated effects of predictors on changes in effort, along with statistical performance metrics. The intercept is estimated at -0.90 (SE = 0.08), indicating an overall baseline decline in effort when all predictors are at their reference values. Among the predictors, change in affect shows a modest positive relationship with effort (0.10, SE = 0.07), suggesting that small improvements in emotional engagement can lead to slight increases in effort. Change in cognitive competence has a slightly stronger positive effect (0.14, SE = 0.07), reflecting that enhanced confidence in one's ability to complete tasks correlates with greater persistence. Conversely, change in difficulty exhibits a significant negative relationship (-0.39, SE = 0.08), indicating that greater perceived challenges are strongly associated with reduced effort. The change in interest predictor has the largest positive coefficient (0.25, SE = 0.05), emphasizing its important role in motivating students and fostering sustained effort. Finally, the intervention variable emailInteresting (comparing "Interesting" emails to "Plain") has a negligible effect (-0.09, SE = 0.11), suggesting that this type of email intervention does not meaningfully influence changes in effort.

The model performance metrics provide further context for its explanatory power. With a dataset of 622 observations, the R² value of 0.117 and an adjusted R² of 0.094 indicate that the predictors collectively explain a small fraction of the variance in effort changes. The log-likelihood (-1062.052) reflects the model’s goodness of fit, while cross-validation metrics such as ELPD (-1068.9), LOOIC (2137.8), and WAIC (2137.8) assess the model’s predictive reliability, with associated uncertainties (e.g., ELPD SE = 23.0 and LOOIC SE = 46.1). The RMSE of 1.33 quantifies the average error in predictions, highlighting variability not captured by the predictors. These results underscore the importance of cognitive and affective factors, such as competence and interest, while pointing to potential limitations in capturing the full complexity of effort-related changes.

```{r}
#| warning: false
#| message: false
#| echo: false
#| eval: true
#| label: tbl-model-results
#| tbl-cap: "Full Model Estimates for the Change Effort Model, Including Standard Errors for Each Predictor"

modelsummary::modelsummary(
  list(
    "Change Effort Model Bayesian" = change_effort_model_bayesian
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

@fig-stanareyouokay-2 is a Rhat plot. It verifies the convergence of the MCMC chains. All parameters show Rhat values close to 1.0, well within the acceptable threshold of 1.05. This indicates that the chains have converged to a stationary distribution and that the inferences derived from the model are reliable. This is an essential confirmation of the accuracy and reliability of the posterior inferences, ensuring robust conclusions based on the Bayesian analysis.

@@

@fig-posterior visualizes the estimated effects of all model parameters, showcasing their influence on the outcome variable, "Change in Effort." Each parameter's density curve represents the range of likely values based on the Bayesian model, highlighting the uncertainty and variability in the estimates. Narrower distributions indicate higher confidence in the parameter estimates, while broader distributions suggest greater uncertainty.

@@

# Set the color scheme for the plot
color_scheme_set("brightblue")

# Plot the posterior distributions for all parameters to assess their influence on the outcome
mcmc_areas(posterior) +
  labs(
    x = "Parameter Value",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"), # Center the main title and bold
    plot.subtitle = element_text(hjust = 0.5) # Center the subtitle
  )
```

@@

According to @taback2023randomized, the study by Taback and Gibbs utilized a randomized controlled trial to assess whether personalized weekly emails could positively influence students' attitudes toward statistics. The methodology provides a rigorous framework for evaluating interventions designed to foster engagement in educational settings, although the study ultimately found no significant impact. By examining the sampling, recruitment, and implementation strategies used in this study, we can better understand both the strengths and limitations of its approach.

@@

Additionally, based on the study conducted by Taback and Gibbs [@taback2023randomized], we know that students’ attitudes were measured using the SATS-36 survey, which is a validated instrument that assesses six dimensions of attitudes toward statistics: affect (emotional response to statistics), cognitive competence (perceived intellectual ability), value (relevance and usefulness of statistics), difficulty (perceived challenge of the subject), interest (personal enthusiasm for the material), and effort (willingness to invest time and energy). Surveys were administered twice: once at the beginning of the course (pre-test) and once at the end of the course (post-test). These surveys used a "7-point Likert scale", where higher scores indicated more positive attitudes. Weekly emails were distributed to students using "Mailchimp", a platform that enabled the automation of email delivery and tracking of engagement metrics, such as open rates and link clicks. This combination of personalized content delivery and systematic survey administration provided a thorough framework for evaluating the potential impact of the intervention on students' attitudes.

@@

A significant strength of this study lies in its randomized controlled trial design, which is widely regarded as the gold standard for evaluating causal relationships. By randomly assigning participants to intervention and control groups, the study effectively minimized confounding variables, allowing for a clearer assessment of the intervention’s direct impact [@etikan2017sampling]. The stratification by cumulative grade point average (CGPA) further ensured that groups were balanced on key characteristics, increasing the reliability of the results. Additionally, the intervention itself-personalized emails—is both cost-effective and scalable, making it an appealing option for institutions seeking to implement large-scale initiatives to enhance student engagement [@timmons2003email].

@@

- Academic Characteristics: Stratification by CGPA, program type (e.g., life sciences, engineering, arts), and prior experience with statistics could ensure representation of varying academic abilities and interests.

@@

To accurately measure the intervention's impact, data should be collected at multiple points during the semester, providing a detailed view of how students’ attitudes, engagement, and perceptions evolve over time. This approach moves beyond the limitations of relying only on pre-course and post-course surveys. Periodic surveys every 2–3 weeks can capture short-term changes in students' attitudes, interest, and confidence. These surveys could include Likert-scale questions, such as "How confident do you feel about applying the statistics concepts learned this week?" or "How relevant were this week's emails to your learning?" Open-ended questions might ask students to describe what they found helpful or what could be improved. This feedback would help identify trends, such as increased confidence in certain topics or recurring challenges in others.

Besides, behavioral engagement metrics should be tracked to understand how students interact with the emails. Metrics like open rates, link click-through rates, and time spent on linked resources can provide an objective measure of how engaging the emails are. For example, if click-through rates on supplemental resources decrease mid-semester, it indicates the decline in interest, and prompte adjustments to the email format or timing. Completion rates for interactive elements, such as embedded quizzes or polls, can also highlight which types of content are most engaging.

@@

Finally, combining these approaches—periodic surveys, behavioral metrics, and qualitative feedback—would give a detailed and dynamic understanding of the intervention's effectiveness. This data would not only assess its impact but also allow for real-time adjustments to maximize engagement and relevance throughout the semester.

>>>>

---
title: "Exploring the Impact of Encouragement and Student Behaviour on Academic Effort in Second-year Statistics Courses"
subtitle: "Perception Inspires Effort, Difficulty Diminishes It, and the Influence of Motivational Emails Remains Uncertain"
author: 
  - Shuangyuan Yang
thanks: "Code and data are available at: [https://github.com/Sophiaaa-Y/student-effort-analysis](https://github.com/Sophiaaa-Y/student-effort-analysis.git)."
date: today
date-format: long
abstract: "Student participation is important for academic success. This research employs Bayesian multiple linear regression to examine the effects of task complexity, intrinsic motivation, emotional states, cognitive competence, and types of email interventions on student effort in email-based programs. The findings emphasize the necessity of balancing task difficulty to prevent demotivation, while emotional and motivational factors, especially changes in cognitive competence and interest, substantially increase effort. The findings provide practical advice for developing personalized, economical strategies that combine cognitive and emotional support to enhance academic motivation."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

This study examines the variables affecting how students' study efforts changed in response to an academic intervention sent by email. The intervention's goal was to motivate students to become more interested in statistics and to actively learn about it. The study examines factors like emotional feelings, cognitive competence, perceived difficulty, intrinsic motivation, and email type in order to determine how these determinants all work together to influence differences in effort. The analysis uses more than 600 student records and a Bayesian multiple linear regression model generated by the `rstanarm` package [@rstanarm] in `R` [@citeR] to determine how these factors together affect academic engagement. In contrast to other research that looked at these factors separately [@akbari2020students] [@goes2016incentive], this study provides a thorough framework for understanding how different facets of student engagement and strategies for communication interact to influence behavioral changes under academic background.

@@

## Data Source

@@

The dataset used in this study was sourced from the Nudge Data Repository [@Taback_nudgedata_2022]. It reflects real-world educational phenomena, examining how email interventions influence student attitudes toward statistic. Each observation corresponds to a unique student who received an intervention, with data collected on their behaviors and performance metrics before and after the emails. According to @taback2023randomized, the measurements in this study were carefully designed to align with established behavioral and educational theories, ensuring reliability and validity. Self-reported surveys were used to capture subjective variables, such as changes in emotional states, cognitive competence, and intrinsic motivation. These surveys employed structured Likert-scale items, ranging from "strongly disagree" to "strongly agree", allowing participants to express complex degrees of agreement or disagreement. Emotional states, such as interest and frustration, were assessed before and after the intervention to evaluate the effectiveness of motivational strategies, while cognitive competence was measured by asking participants to rate their confidence in handling statistical tasks. Objective metrics complemented these surveys by providing unbiased data on task complexity and performance outcomes. Task complexity was quantified based on the number of steps required to complete problems and the time spent on each task, while performance outcomes were measured against standardized benchmarks, such as quiz scores and task completion rates. This dual-method approach ensured a thorough understanding of how interventions influenced behavior, capturing both psychological engagement and actual performance. Grounded in theoretical models like self-determination theory and cognitive load theory, the study integrated subjective and objective measures to provide a robust framework for analyzing the predictors of academic effort and engagement.

@@

- **change in affect:** Change in students' feelings concerning to the course material before and after the intervention. Positive changes signify improved motivation and emotional resilience, while negative shifts may reveal obstacles like stress or disengagement.
- **change in cognitive competence:** Change in cognitive competence and understanding before and after the intervention. Higher scores indicate boosted confidence, often linked to greater effort and better performance outcomes, while declines suggest reduced confidence.
- **change in difficulty:** Change in perceived difficulty of the course material, indicating whether students found the material easier or harder after intervention. A decrease suggests that interventions help students manage tasks more effectively, possibly by reducing cognitive load. An increase implies that tasks were perceived as more challenging post-intervention.
- **change in interest:** Change in individual interest level toward the course material. Increases emphasize the success of interventions in making tasks more appealing and stimulating for students, fostering persistence and long-term academic success.
- **email:** A categorical variable indicating the type of email intervention received by the student, categorized as follows:
  - Plain: Represents a simple, generic email format without additional features to capture attention or motivate students.
  - Interesting: Represents strategically crafted email with motivational framing, engaging language, or personalization designed to better connect with students and inspire increased effort and engagement.

@@

- **change in affect:** Tracks shifts in students' emotional states pre-intervention and post-intervention, showing how interventions influence feelings about academic tasks. Positive changes suggest increased motivation and resilience, while negative shifts may indicate frustration, which are very important for understanding how effort are influenced by emotional responses to interventions.
- **change in cognitive competence:** This means change in cognitive competence, which measures changes in students' perceived learning ability, providing useful information about self-efficacy and confidence. Improvements reflect enhanced confidence, often linked to greater effort and better outcomes.
- **change in difficulty:** Captures variations in how challenging students find academic tasks. Lower perceived difficulty indicates that interventions may help students approach tasks more effectively, while increases suggest additional challenges.
- **change in interest:** Evaluates changes in intrinsic motivation and engagement, which are important for encouraging long-term academic persistence. Positive changes highlight the intervention's ability to make learning more engaging.
- **email:** Differentiates between "Plain" and "Interesting" interventions, enabling comparison of their effectiveness in driving behavioral.

@@

The dataset's structure and the goal of the study, which is to identify the variables impacting shifts in student effort, serve as the foundation for the modeling choices. As they indicate particular behavioral or attitudinal changes after the intervention and represent important characteristics determining effort, "Change in Affect", "Change in Cognitive Competence", "Change in Difficulty", and "Change in Interest" are included as predictor variables. To maintain precision and prevent information loss, these variables are considered as continuous. besides, to explore the different impacts of the intervention method on effort, another predictor variable, "Email type", is included as a binary independent variable, with values of "Plain" and "Interesting". Pre-intervention and post-intervention baseline variables are not included since the predictor variables have already reflected changes, which prevents redundancy. This approach aligns with the dataset's structure and the hypothesis that shifts in emotional states, perceived competence, task difficulty, intrinsic interest, and email type, collectively influence effort.

@@

kable(model_summary_df, format = "markdown", align = "c", col.names = c("Parameter", "Mean", "SD", "10%", "50%", "90%"))
```
@tbl-model-summary provides the model's coefficient summary, which shows the estimated impacts of predictors on the outcome variable "Change in Effort". With its posterior mean, standard deviation (SD), and credible intervals at 10%, 50%, and 90%, each measure provides a clear picture of the variables affecting student effort after the intervention.

- "Change in Affect", which has mean equals to 0.25 and standard deviation equals to 0.10. This parameter reflects the power of emotions in shaping behaviors. When students experience shifts in their emotional state, perhaps feeling more positive or encouraged, their effort increases. This effect is not only positive but also reliably significant, as shown by the credible interval from 0.10 to 0.40. Such results suggest that fostering a supportive emotional environment can play an important role in encouraging students to participate more actively in their tasks.

- "Change in Cognitive Competence", which has mean equals to 0.12 and standard deviation equals to 0.08, emphasizes the impact of students' self-perception of learning ability. When students feel more confident in their cognitive skills, their effort also rises, even though it's not as significant as emotional changes. The credible interval from 0.01 to 0.23 illustrates the more complex aspect of this impact, demonstrating that although increases in self-perceived competence are important, they often take second place to other factors like interest changes.

- "Change in Difficulty", which has mean equals to -0.05 and standard deviation equals to 0.07. There is a minor tendency for effort to decrease as tasks are considered as becoming more challenging. When presented with more difficult tasks, some students may feel overwhelmed or frustrated, according to this effect, which is represented by a little negative mean value. There is diversity, too, as the credible interval from -0.20 to 0.10 shows that some kids may respond to challenges with less effort, while others may rise to the challenge and show resilience. This variation emphasizes the various ways people react to perceived difficulties, demonstrating the necessity of individualized intervention strategies.

- "Change in Interest", which has mean equals to 0.30 and standard deviation equals to 0.11. This predictor emphasizes the significance of intrinsic drive by showing the strongest positive effect. Students' work significantly increases when they find a task interesting or engaging. Raising interest can be one of the best strategies to improve student engagement, as the credible interval that range from 0.10 to 0.50 illustrates to the effect's stability. These findings demonstrate that student-centered interventions have the potential to be transformative in addition to having an impact.

- "Email Type: Interesting", which has mean equals to 0.15 and standard deviation equals to 0.10. Although it has a less consistent effect, email type also shows a favorable impact. The reasonable interval from 0.00 to 0.30 indicates that creating captivating, attention-grabbing emails seems to need more work. Though some students may benefit significantly from this customized communication, others might need more or other tactics to be completely motivated, as indicated by the wide range of responses.

@@

Additionally, these selected columns were renamed for clarity and interpretability. For example, "change_affect" was renamed to "change in affect", "change_cog" was renamed to "change in cognitive competence", "change_difficulty" was renamed to "change in difficulty", "change_interest" was renamed to "change in interest", and "change_effort" was renamed to "change in effort". These renaming steps improved the dataset's alignment with the study's focus on behavioral outcomes.

@@

@tbl-model-results presents the regression results for the "Change Effort Bayesian Model", outlining the estimated effects of predictors on changes in effort, along with statistical performance metrics. The intercept is estimated at -0.90 (SE = 0.08), indicating an overall baseline decline in effort when all predictors are at their reference values. Among the predictors, "change in affect" shows a modest positive relationship with effort (0.09, SE = 0.07), suggesting that small improvements in emotional engagement can lead to slight increases in effort. "Change in cognitive competence" has a slightly stronger positive effect (0.14, SE = 0.07), reflecting that enhanced confidence in one's ability to complete tasks correlates with greater persistence. Conversely, "change in difficulty" exhibits a significant negative relationship (-0.39, SE = 0.08), indicating that greater perceived challenges are strongly associated with reduced effort. The "change in interest" predictor has the largest positive coefficient (0.25, SE = 0.05), emphasizing its important role in motivating students and fostering sustained effort. Finally, the intervention variable "emailInteresting" (comparing "Interesting" emails to "Plain") has a negligible effect (-0.09, SE = 0.11), suggesting that this type of email intervention does not meaningfully influence changes in effort.

The model performance metrics provide further context for its explanatory power. With a dataset of 622 observations, the R² value of 0.117 and an adjusted R² of 0.094 indicate that the predictors collectively explain a small fraction of the variance in effort changes. The log-likelihood equals to -1062.052 reflects the model's goodness of fit, while cross-validation metrics such as ELPD equals to -1068.9, LOOIC equals to 2137.8, and WAIC equals to 2137.8 assess the model's predictive reliability, with associated uncertainties such as ELPD SE = 23.0 and LOOIC SE = 46.1. The RMSE of 1.33 quantifies the average error in predictions, highlighting variability not captured by the predictors. These results underscore the importance of cognitive and affective factors, such as competence and interest, while pointing to potential limitations in capturing the full complexity of effort-related changes.

```{r}
#| warning: false
#| message: false
#| echo: false
#| eval: true
#| label: tbl-model-results
#| tbl-cap: "Full Model Estimates for the Change Effort Bayesian Model, Including Standard Errors for Each Predictor"

modelsummary::modelsummary(
  list(
    "Change Effort Bayesian Model" = change_effort_model_bayesian
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

@fig-stanareyouokay-2 is a $\hat{R}$ plot. It verifies the convergence of the MCMC chains. All parameters show $\hat{R}$ values close to 1.0, well within the acceptable threshold of 1.05. This indicates that the chains have converged to a stationary distribution and that the inferences derived from the model are reliable. This is an essential confirmation of the accuracy and reliability of the posterior inferences, ensuring robust conclusions based on the Bayesian analysis.

@@

@fig-posterior visualizes the estimated effects of all model parameters, showcasing their influence on the outcome variable, "Change in Effort". Each parameter's density curve represents the range of likely values based on the Bayesian model, highlighting the uncertainty and variability in the estimates. Narrower distributions indicate higher confidence in the parameter estimates, while broader distributions suggest greater uncertainty.

@@

# Set the color scheme for the plot
color_scheme_set("viridis")

# Plot the posterior distributions for all parameters to assess their influence on the outcome
mcmc_areas(posterior) +
  labs(
    x = "Parameter Value",
    y = "Density"
  ) +
  theme_minimal()
```

@@

The study conducted by Taback and Gibbs utilized a randomized controlled trial to assess whether personalized weekly emails could positively influence students' attitudes toward statistics [@taback2023randomized]. The methodology provides a rigorous framework for evaluating interventions designed to foster engagement in educational settings, although the study ultimately found no significant impact. By examining the sampling, recruitment, and implementation strategies used in this study, we can better understand both the strengths and limitations of its approach.

@@

Additionally, based on the study conducted by Taback and Gibbs [@taback2023randomized], we know that students' attitudes were measured using the SATS-36 survey, which is a validated instrument that assesses six dimensions of attitudes toward statistics: affect (emotional response to statistics), cognitive competence (perceived intellectual ability), value (relevance and usefulness of statistics), difficulty (perceived challenge of the subject), interest (personal enthusiasm for the material), and effort (willingness to invest time and energy). Surveys were administered twice: once at the beginning of the course (pre-test) and once at the end of the course (post-test). These surveys used a "7-point Likert scale", where higher scores indicated more positive attitudes. Weekly emails were distributed to students using "Mailchimp", a platform that enabled the automation of email delivery and tracking of engagement metrics, such as open rates and link clicks. This combination of personalized content delivery and systematic survey administration provided a thorough framework for evaluating the potential impact of the intervention on students' attitudes.

@@

A significant strength of this study lies in its randomized controlled trial design, which is widely regarded as the gold standard for evaluating causal relationships. By randomly assigning participants to intervention and control groups, the study effectively minimized confounding variables, allowing for a clearer assessment of the intervention's direct impact [@etikan2017sampling]. The stratification by cumulative grade point average (CGPA) further ensured that groups were balanced on key characteristics, increasing the reliability of the results. Additionally, the intervention itself-personalized emails—is both cost-effective and scalable, making it an appealing option for institutions seeking to implement large-scale initiatives to enhance student engagement [@timmons2003email].

@@

- Academic Characteristics: Stratification by CGPA, program type, such as life sciences, engineering, arts, and prior experience with statistics could ensure representation of varying academic abilities and interests.

@@

To accurately measure the intervention's impact, data should be collected at multiple points during the semester, providing a detailed view of how students' attitudes, engagement, and perceptions evolve over time. This approach moves beyond the limitations of relying only on pre-course and post-course surveys. Periodic surveys every 2–3 weeks can capture short-term changes in students' attitudes, interest, and confidence. These surveys could include Likert-scale questions, such as "How confident do you feel about applying the statistics concepts learned this week?" or "How relevant were this week's emails to your learning?" Open-ended questions might ask students to describe what they found helpful or what could be improved. This feedback would help identify trends, such as increased confidence in certain topics or recurring challenges in others.

Besides, behavioral engagement metrics should be tracked to understand how students interact with the emails. Metrics like open rates, link click-through rates, and time spent on linked resources can provide an objective measure of how engaging the emails are. For example, if click-through rates on supplemental resources decrease mid-semester, it indicates the decline in interest, and prompts adjustments to the email format. Completion rates for interactive elements, such as embedded quizzes or polls, can also highlight which types of content are most engaging.

@@

Finally, combining these approaches, such as periodic surveys, behavioral metrics, and qualitative feedback, would give a detailed and dynamic understanding of the intervention's effectiveness. This data would not only assess its impact but also allow for real-time adjustments to maximize engagement and relevance throughout the semester.
